# The backbone: The relational database

The approach I take in this chapter is to start by illustrating the kinds of things that a relational database platform allows me to do.
Initially, I will gloss over the messy details of setting up a database, getting data into it, etc.
But rest assured that if you like what you see, it is fairly straightforward to set up the infrastructure I describe.^[Much of the code I use to maintain my platform, along with instructions for using it, is available on my GitHub [page](https://github.com/iangow).]

In addition, I will often illustrate ideas using code that I only fully explain later in the book.
For example, I find R packages such as `dplyr` to be particularly slick
```{r, include = FALSE}
Sys.setenv(PGDATABASE="crsp", PGHOST="iangow.me")
```

## Handling large-ish data sets

One nice feature of a relational database is that it makes it easy to access data from 
large-ish data sets.

Suppose I was interested in the stock returns of Amazon.^[This is admittedly not a particularly realistic *research* problem, but many of the benefits described here do accrue in real research applications.]

The CRSP daily stock file (denoted as `crsp.dsf` on the WRDS system) contains about 18 GB of data in about 90 million rows.
While someone dealing with the log files of a busy website or managing the geospatial data of Uber would chuckle at the idea of this being "big data", the reality is that 18GB is a pain to work with in a number of situations:

- When the data aren't local. Downloading 18GB over the internet is time-consuming with even the fastest broadband connections.
- When you have limited RAM. For a typical laptop in 2017, loading in 18GB into memory will exhaust available RAM. 
- When you're only interested in a small sliver of the data, such in our hypothetical "Amazon stock returns" case.

So how would I access these data? First, I would establish a connection to my database:

```{r}
library(dplyr, warn.conflicts = FALSE)
pg <- src_postgres()
```

Here I am using R (and, in particular, the package `dplyr`). 
The first line loads the package.^[You may have to `install.packages("dplyr")` if you do not have it installed already.] 
The second line connects to my database.^[I have omitted connection information here. As I will discuss later, I think it is good practice to supply that elsewhere, which is what I have done here.]

The next thing I need to do is to identify the `permno` (CRSP's firm identifier), which I can get by looking at the `crsp.stocknames` table.
```{r}
stocknames <- tbl(pg, sql("SELECT * FROM crsp.stocknames"))

stocknames %>% 
  filter(comnam %~*% 'AMAZON') %>%
  select(permno, comnam) %>%
  collect()
```
The first line of the code creates a variable `stocknames` that I can use to refer to the underlying data table `crsp.stocknames` in my database.^[This syntax requires the use of a full SQL statement, `SELECT * FROM crsp.stocknames`, which isn't particularly elegant, but also isn't particularly problematic.]

Here I am using the `%>%` "pipe" operator, which is comes from the `magrittr` package and is re-exported by `dplyr`.
The easiest way to understand this in the context of `dplyr` is that thing to the left of `%>%` is typically the object of the thing to the right of `%>%`, which is typically a verb.
So we take stocknames, and `filter` to retain only observations where the company name (`comnam`) matches `AMAZON`.^[`%~%` is PostgreSQL's regular expression matching operator. More will be said on all this later on. For now, just accept that it's doing some kind of string matching.]
Now that I know that the PERMNO for Amazon is `84788`, I can go to the daily stock file (`crsp.dsf`) to get stock returns.^[An alternative approach would involve joining the query above with the one below. But I assume that it is necessary to visually inspect the query above to verify that we have the right "Amazon" here.]

The result from that `filter` operation is another (smaller) data set that becomes the object for the next verb: `select`.
Like its SQL counterpart (typically written `SELECT` in SQL circles), `select` chooses the columns of the data set that are to be retained.
For this query, we want the `permno` and, to check we have the right company, the company's name (`comnam`).

```{r}
dsf <- tbl(pg, sql("SELECT * FROM crsp.dsf"))

amzn_rets <-
  dsf %>% 
  filter(permno == 84788L) %>%
  select(permno, date, ret, prc) %>%
  mutate(prc = abs(prc)) %>%
  collect()

amzn_rets
```

Now, we set up a variable to refer to `crsp.dsf`, `filter` to get data just for Amazon, `select` the variables of interest, then `mutate` the stock price variable (`prc`) to correct negative values, before using `collect` to bring the data from the database into R.^[Stock prices should be positive, but [André de Souza](http://pages.stern.nyu.edu/~adesouza/sasfinphd/index/node17.html) of NYU points out that "closing prices in the `crsp.msf` and `crsp.dsf` files are negative if the price is not the price from an actual trade, but the average of bid-ask spreads. I always set `prc = abs(prc)` before I do anything else with price."]

Note that this all took about a tenth of a second. 
So the fact that the data are stored in an 18GB table on a different computer was not at all problematic.
PostgreSQL is able to grab the data of interest quickly and send (over the internet) just what I need.
The resulting data is a mere `r format(object.size(amzn_rets), units = "KB")`.

Now that I have the data, suppose that I wanted to plot cumulative stock performance.
Many researchers would take the data above and dump to a file that could be opened in Excel, open that file in Excel, use the mouse to select the data of interest, and then make a plot.
That plot could be copied and pasted into, say, a Word document.
But here we have taken several manual steps of the kind I argued against in Chapter 2.

How else might we do this? Well, we could use the excellent `ggplot2` package to make a plot following on from our steps so far. Here is the code to do so:

```{r}
library(ggplot2)
amzn_rets %>% 
  mutate(cumret = cumprod(1+coalesce(ret, 0)) - 1) %>%
  ggplot(aes(x = date, y = cumret)) +
  geom_line() +
  xlab("Cumulative returns") +
  ggtitle("Amazon's stock performance") + 
  theme(plot.title = element_text(hjust = 0.5))
```
  
